Geometric variational problems

### 2020-11-14 NSF2020/Description/background.tex
fixed bibiography and touched up SA3
prelim done

\noindent
{\bf Collaborative Research: Mathematical modeling and simulations of
self-assembling amphiphilic particles in a solvent} \\
%{\bf {Collaborative Research: Mathematical modeling and coarse-grained simulations of self-assembly of amphiphilic Janus particles in a solvent}}\\
{\em Rolf Ryham (lead PI, Fordham University), Bryan Quaife (PI,
FSU), and Y.-N. Young (PI, NJIT)}
\section{Background}
\label{sec:background}
The goal of this collaborative proposal is to use mathematical modeling
and numerical simulations to investigate the dynamic self-assembly of
amphiphilic particles interacting via hydrophobic forces
in a solvent. Amphiphilic particles (such as lipid molecules) possess
both hydrophobic and hydrophilic structures. In a viscous solvent they
self-assemble into meso-/macroscopic structures (such as micelles and
bilayers of lipids) to shield their hydrophobic parts from contact with
the solvent molecules (water).
%
%\subsection{Hydrophobic Forces}
%\label{sec:hydrophobicforce}
%
Such self-assembly of amphiphiles via hydrophobic forces is ubiquitous in biology and biophysics \cite{Israelachvili1954},
%The hydrophobic force is a ubiquitous molecular interaction in biology \cite{Israelachvili1954}, 
and has been a major source of nonspecific interactions between
nanoparticles in soft matter
\cite{Sanchez-IglesiasEtAl2012_ACSNano,AltantzisEtAl2013_PSC,XieYangLuEtAl2020_COCIS}. 
%The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles so to design smart materials of desirable properties by
%tuning the geometry and properties of the amphiphilic particles.

@@

The time integration of
particle configurations goes as follows: \textbf{(i)} solve the
BVP~\eqref{SL} for the screened Laplace equation, \textbf{(ii)}
determine the rigid body forces and torques~\eqref{forceandtorque},
\textbf{(iii)} solve the Stokes mobility problem~\eqref{eq:stokes} for
rigid body motion, \textbf{(iv)} update the particle configuration.
Numerical challenges and how these challenges are overcome are described
in \S\ref{subsec:specific_aim_2} and \S\ref{subsec:specific_aim_3}.

>>>>

\noindent
{\bf Collaborative Research: Mathematical modeling and simulation of
self-assembling amphiphilic particles in solvent} \\
%{\bf {Collaborative Research: Mathematical modeling and coarse-grained simulations of self-assembly of amphiphilic Janus particles in a solvent}}\\
{\em Rolf Ryham (lead PI, Fordham University), Bryan Quaife (PI,
FSU), and Y.-N. Young (PI, NJIT)}
\section{Background}
\label{sec:background}
The goal of this collaborative proposal is to use mathematical modeling
and numerical simulations to investigate the dynamic self-assembly of
amphiphilic particles interacting via hydrophobic forces
in a solvent. Amphiphilic particles (such as lipid molecules) possess
both hydrophobic and hydrophilic structures. In a viscous solvent they
self-assemble into meso-/macroscopic structures (such as micelles and
bilayers of lipids) to shield their hydrophobic parts from contact with
the solvent molecules (water).
%
%\subsection{Hydrophobic Forces}
%\label{sec:hydrophobicforce}
%
Such self-assembly of amphiphiles via hydrophobic forces is ubiquitous in biology and biophysics \cite{Israelachvili1954},
%The hydrophobic force is a ubiquitous molecular interaction in biology \cite{Israelachvili1954}, 
and has been a major source of nonspecific interactions between
nanoparticles in soft matter
\cite{Sanchez-IglesiasEtAl2012_ACSNano,AltantzisEtAl2013_PSC,XieYangLuEtAl2020_COCIS}. 
%The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles so to design smart materials of desirable properties by
%tuning the geometry and properties of the amphiphilic particles.

@@

The time integration of
particle configurations goes as follows: \textbf{(i)} solve the
BVP~\eqref{SL} for the screened Laplace equation, \textbf{(ii)}
determine the rigid body forces and torques~\eqref{forceandtorque},
\textbf{(iii)} solve the Stokes mobility problem~\eqref{eq:stokes} for
rigid body motion, \textbf{(iv)} update the particle configuration.
Numerical challenges and how these challenges are overcome are described
in \S\ref{subsec:specific_aim_2}.

### 2020-11-14 NSF2020/Description/preliminarywork.tex
RJR
prelim done

\section{Preliminary work}
\label{sec:preliminary_work}
The HAP project was initiated in 2017 by PIs RR and YNY.
The lead author on our first paper (\cite{Fu2018_SIAM}) was Szu-Pei Fu,
who was PI YNS's PhD student and is currently a postdoc with PI RR.
The collaboration involved theoretical work and simulation. The theoretical
work included model development, proving the first variation formula
\eqref{stress} and force-free law, and an energy-based  uniqueness
theorem for the exterior domain problem \eqref{SL}.
Coauthors Andreas Kl\"ockner (Computer Science, UIUC, NSF CAREER \#1654756)
and his PhD student Matt Wala provided their Quadrature by Expansion, or QBX
algorithm and Fu ran simulations on NJIT clusters.

@@

\begin{wraptable}[10]{r}{0.55\textwidth}
\begin{center}
  \begin{tabular}{|l|r|r|p{0.1\textwidth}|}
    \hline
          & HAP                     & exper. value          & ref.\\
    \hline 
    $\KB$ & $11 \pm 2.5$ \kBT       & 10 \kBT               & \cite{Naetal15,VeBrPa15,NAGLE2000159,PhysRevLett.113.248102}\\
    \hline 
    $\KA$ & $34$ \kBT \; nm$^{-2}$  & 30--40 \kBT nm$^{-2}$ & \cite{Nagle17, Nagle17-2}\\
    \hline 
    $\KTH$ & $12$ \kBT \; nm$^{-2}$ & 10 \kBT nm$^{-2}$     & \cite{KUZMIN2005, KoNa15} \\ \hline
  \end{tabular}
\end{center}
\caption{\label{tab:moduli} \footnotesize Comparision of values elastic
  moduli from the experimental literature and values derived by HAP
  simulation.} 
\end{wraptable}
We performed three types of simulations. The first measured bending by loading
a planar bilayer. The second used a harmonic bond to dilate a circular bilayer.
The third measured tilt using a decay equation from  \cite{KUZMIN2005}.
These simulations isolated the deformations of \eqref{ansatz3}, enabling
us to read off elastic moduli from simulation data. The results agreed remarkably 
well with the values reported in the experimental literature, 
Table~\ref{tab:moduli}. The agreement is underscored by the fact
that the two, main HAP parameters $\gamma$ and $\rho$ were assigned
physical values from the outset, rather than being tuned.

The PIs BQ, RR and YNY initiated a collaboration in summer, 2020 to develop
the hydrodynamic model of \S \ref{subsec:specific_aim_3}
To date, we have written solvers for the two-dimensional boundary integral formulatons
of \eqref{SL} and \eqref{eq:stokes}. The solvers gives third-order spatial accuracy
up to the boundary for $u$ and $\mathbf{u}$. The specialized reciprocal identity \ref{eq:reciprocal}
gives spectral accuracy for force and torque (Table \ref{tab:spectral_force}}).
\begin{wraptable}[10]{l}{0.45\textwidth}
\begin{center}
  \begin{tabular}{|l|l|l|l|}
 $n$   & $\max_{i} \|\mathbf{F}_i-\mathbf{F}_i^{\text{exact}}\|$   &  $\max_{i} \|\mathbf{G}_i-\mathbf{G}_i^{\text{exact}}\|$   \\
 32   &  81.4933     &   70.3970\\       
  64  &   4.9328     &    3.1928 \\
 128  &   0.0125     &    0.0030 \\
256   &   0.0000765   &   0.00003389        
  \end{tabular}
\end{center}
\caption{\label{tab:spectral_force} 
\footnotesize 
Numerical error for force and torque as a function of number of grid points $n$ per particle.
The data is for $N = 5$ particles characteristic diamater 2; two of the particles are nearly touching 
at a distance 0.02.} 
\end{wraptable}

We have simulated two-dimensional vesicles in shear flow 
$\mathbf{u}_{\infty} = \dot{\gamma} x_2 \mathbf{i}_1$ with shear rate $\dot{gamma}$. 
The particle-based system mimiks the behavior observed in continuous vesicles
\cite{torres-sanchez_millan_arroyo_2019, mahapatra_saintillan_rangamani_2020, 
Steigmann99, C6SM02452A}. The basic pattern is that at moderate shear rates,
the initially circular vesicle bilayer elongates and approaches a steadily 
tank-treading ellipse. Analyzing any underlying constituative laws, the particles
reorganize so that there is an initial decrease in vesicle area that plateaus 
to a steady state area, respectively increase in arc length. This suggests that
the particle system behaves as a semipermeable vesicle \cite{323e9a2f0c58487ea82518d7a1f96485,YAO2017728}.
Our preliminary results already showed that the particle-based bilayers have the same
large area modulus as real lipid bilayer membranes, and permit a small amount of stretching. 
For large shear rates, the vesicle ruptures and flow carries away segmented membrane patches.

The above considerations justify using the particle-based model to study material properties 
as outline in \S \ref{subsec:specific_aim_1}. Furthermore, the particle system allows
for intermonolayer slip~\cite{SHKULIPA2005823, ShkulipaThesis}.
Intermonolayer slip enters zero-thickness membrane models as a velocity
jump boundary condition \cite{schwalbe_vlahovska_miksis_2010}, for example, 
but in the present model is a consequence of monolayer independence.

>>>>

\begin{wraptable}[9]{r}{0.55\textwidth}
\centering{
  \begin{tabular}{|l|r|r|p{0.1\textwidth}|}
    \hline
          & HAP                     & experiment          & ref.\\
    \hline 
    $\KB$ & $11 \pm 2.5$ \kBT       & 10 \kBT               & \cite{Naetal15,VeBrPa15,NAGLE2000159,PhysRevLett.113.248102}\\
    \hline 
    $\KA$ & $34$ \kBT \; nm$^{-2}$  & 30--40 \kBT nm$^{-2}$ & \cite{Nagle17, Nagle17-2}\\
    \hline 
    $\KTH$ & $12$ \kBT \; nm$^{-2}$ & 10 \kBT nm$^{-2}$     & \cite{KUZMIN2005, KoNa15} \\ \hline
  \end{tabular}
}
\caption{\label{tab:moduli} \footnotesize Comparision of values elastic
  moduli from the experimental literature and values derived by HAP
  simulation.} 
\end{wraptable}
\section{Preliminary work}
\label{sec:preliminary_work}
The HAP project was initiated in 2017 by PIs RR and YNY.
The lead author on our first paper (\cite{Fu2018_SIAM}) was Szu-Pei Fu,
who was PI YNS's PhD student and is currently a postdoc with PI RR.
The collaboration involved theoretical work and simulation. The theoretical
work included model development, proving the first variation formula
\eqref{stress} and force-free law, and an energy-based  uniqueness
theorem for the exterior domain problem \eqref{SL}.
Coauthors Andreas Kl\"ockner (Computer Science, UIUC, NSF CAREER \#1654756)
and his PhD student Matt Wala provided their Quadrature by Expansion, or QBX
algorithm and Fu ran simulations on NJIT clusters.

@@

We performed three types of simulations. The first measured bending by loading
a partially clamped planar bilayer. The second used a harmonic bond to dilate a circular bilayer.
The third measured tilt using a decay equation from  \cite{KUZMIN2005}.
These simulations isolated three of the five deformations of \eqref{ansatz3}, enabling
us to read off elastic moduli from simulation data. The results agreed remarkably 
well with the values reported in the experimental literature 
(Table~\ref{tab:moduli}). The agreement is underscored by the fact
that the two, main HAP parameters $\gamma$ and $\rho$ were assigned
physical values from the outset, rather than being tuned so as to fit data.

The PIs BQ, RR and YNY initiated a collaboration in summer, 2020 to develop
the hydrodynamic model of \S \ref{subsec:specific_aim_3}
To date, we have written solvers for the two-dimensional boundary integral formulatons
of \eqref{SL} and \eqref{eq:stokes}. The solvers gives third-order spatial accuracy
up to the boundary for $u$ and $\mathbf{u}$. The specialized reciprocal identity \ref{eq:reciprocal}
gives spectral accuracy for force and torque (Table \ref{tab:spectral_force}).
\begin{wraptable}[12]{l}{0.40\textwidth}
\centering{
  \begin{tabular}{|l|l|l|l|}
    $n$   & (rel. err.)$_F$
    &  (rel. err.)$_G$ \\
    \hline
 32   &   4.07466     &   1.56437\\
 64   &   0.24664     &   0.03476\\
 128  &   0.00063     &   0.00006\\
  \end{tabular}}
\caption{\label{tab:spectral_force} 
\footnotesize 
Relative numerical errors
(rel. err.)$_F = \max_{i} \|\mathbf{F}_i-\mathbf{F}_i^{\text{exact}}\|/\|\mathbf{F}_i^{\text{exact}}\|$
and 
(rel. err.)$_G = \max_{i} \|\mathbf{G}_i-\mathbf{G}_i^{\text{exact}}\|/\|\mathbf{G}_i^{\text{exact}}\|$
for force and torque respectively as a function of number of grid points $n$ per particle.
The data is for $N = 5$ particles; two of the particles are nearly touching 
at a distance 1 \% of particle diameter.} 
\end{wraptable}

We have simulated two-dimensional vesicles in shear flow 
$\mathbf{u}_{\infty} = \dot{\gamma} x_2 \mathbf{i}_1$ with shear rate $\dot{\gamma}$.
(The symbol shear rate is unrelated to interfacial tension.)
The particle-based system mimiks the behavior observed in continuous vesicles
\cite{torres-sanchez_millan_arroyo_2019, mahapatra_saintillan_rangamani_2020, 
  Steigmann99, C6SM02452A}.
The basic pattern is that the vesicle bilayer approaches a steadily 
tank-treading ellipse (Figure \ref{fig:tank_treading}C).
We therefore analyzed the simulation data in the context of
theoretical work on tank-treading \cite{Finken2008}. 
Jeffrey's equation predicts a tank-treading frequency $\tfrac{1}{2}\dot{\gamma}$
(dashed line Figure \ref{fig:tank_treading}B).
In the particle simulation, the tank-treading frequency corresponds to the mean
of the particle angular velocities, and these approach a value within 7 \% of
the theoretical prediction (dotted and solid curves in Figure \ref{fig:tank_treading}B).
In the low viscosity-contrast regime of our simulation, Jeffrey's equation predicts
a decrease in the ellipse's inclination angle $\Theta$
and our particle-based simulation follows this prediction
(Figure \ref{fig:tank_treading}D). The higher frequency oscillation comes
from particle reorganization.

Analyzing for any underlying constituative laws, the particles
reorganize so that there is an initial increase and then plateauing of vesicle circumference.
As is the case with real lipid bilayer membranes, the particle collection permits a small amount of stretching. 
The theoretical calculations give an increase of length by 3.6 \% which
agrees with our measured increase in length. The simulations also suggest
that the particle system behaves as a semipermeable vesicle \cite{323e9a2f0c58487ea82518d7a1f96485,YAO2017728}.

\begin{wrapfigure}[13]{r}{0.3\textwidth}
  \centering{\includegraphics[width=0.3\textwidth]{Figures/PW_fig5.pdf}}
    \caption{\label{fig:flow_rupture} \footnotesize
      Rupture of a two-dimensional vesicle for large flow rates.}
\end{wrapfigure}
The above considerations further justify using the particle-based model to study material properties 
as outlined in \S \ref{subsec:specific_aim_1}.
Furthermore, the particle system allows
for intermonolayer slip~\cite{SHKULIPA2005823, ShkulipaThesis}.
In Figure \ref{fig:tank_treading}B, the inner angular velocity is slightly
less than the outer angular velocity. 
Intermonolayer slip enters zero-thickness membrane models as a velocity
jump boundary condition \cite{schwalbe_vlahovska_miksis_2010}, for example, 
but in the present model, it is a consequence of monolayer independence.
For large shear rates, viscous forces exceed the hydrophobic attraction and the vesicle ruptures;
the flow carries away segmented membrane patches (Figure \ref{fig:flow_rupture}).

\begin{figure}%[13]{r}{0.61\textwidth}
  \centering{\includegraphics[width=0.61\textwidth]{Figures/TankTreading.pdf}}
    \caption{\label{fig:tank_treading} \footnotesize
    A two-dimemsional particle-based vesicle under shear flow. 
    Panel A shows the particles, water activity, and background flow direction.
    In Panel C, the trace of the particle centers forms two ellipses with well-defined
    inclination angle $\Theta$.
    Panel B plots the mean angular velocities for the outer leaflet and the inner respectively.
    Panel D shows the evolution of the inclination angle. 
    Panels B and D share the same time axis.}
\end{figure}
